---
title       : Data Product 
subtitle    : Galton's Height Data
author      : Claudio Seidi Takamiya
job         :
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction 
This work is based on Galton's Height Data.


* The Galton's Heigth Data can be found in HistData library
* There is a more detailed data inclunding child gender. The dataset is 
GaltonFamilies
* This work checks the influence of the child gender on its height.

The code below load the GaltonFamilies dataset.

```{r  warning=FALSE}
library(HistData)
data(GaltonFamilies)
```

--- .class #id 

## Exploratory Analysis
```{r echo=FALSE}

gf <- GaltonFamilies
gf$aux <- paste(gf[,2],",",gf[,3])
gfMale <- gf[gf$gender=="male", ] 
gfFemale <- gf[gf$gender=="female", ] 


gf.df <- data.frame(x = gf[, 2], y = gf[, 3], z = gf[, 8])
gfMale.df <- data.frame(x = gfMale[, 2], y = gfMale[, 3], z = gfMale[, 8])
gfFemale.df <- data.frame(x = gfFemale[, 2], y = gfFemale[, 3], z = gfFemale[, 8])

gf.loess = loess(z ~ x * y, data=gf.df, degree=2, span=0.5)
gfMale.loess = loess(z ~ x * y, data = gfMale.df, degree=2, span=0.5)
gfFemale.loess = loess(z ~ x * y, data = gfFemale.df, degree=2, span=0.5)

gf.fit = expand.grid(list(x=seq(60,80,0.5), y=seq(60,80,0.5)))
z <- predict(gf.loess, newdata=gf.fit)
zMale <- predict(gfMale.loess, newdata=gf.fit)
zFemale <- predict(gfFemale.loess, newdata=gf.fit)

gf.resid <- resid(gf.loess)
gfMale.resid <- resid(gfMale.loess)
gfFemale.resid <- resid(gfFemale.loess)


jet.colors <- colorRampPalette(c("blue", "green"))
nbcol <- 100;
color <- jet.colors(nbcol)
nrz <- nrow(z)
ncz <- ncol(z)
zfacet <- z[-1,-1] + z[-1,-ncz] + z[-nrz, -ncz]
facetcol <- cut(zfacet, nbcol)



jet.colors2 <- colorRampPalette(c("darkblue", "yellow"))
nbcol2 <- 100;
color2 <- jet.colors2(nbcol2)
nrz2 <- nrow(zMale)
ncz2 <- ncol(zMale)
zfacet2 <- zMale[-1,-1] + zMale[-1,-ncz2] + zMale[-nrz2, -ncz2]
facetcol2 <- cut(zfacet2, nbcol2)


jet.colors3 <- colorRampPalette(c("darkred", "purple"))
nbcol3 <- 100;
color3 <- jet.colors3(nbcol3)
nrz3 <- nrow(zFemale)
ncz3 <- ncol(zFemale)
zfacet3 <- zFemale[-1,-1] + zFemale[-1,-ncz3] + zFemale[-nrz3, -ncz3]
facetcol3 <- cut(zfacet3, nbcol3)
```

```{r echo=FALSE,fig.width=13, fig.height=6}
par(bg="#E0E0E0", mfrow=c(1,3))
persp(seq(60,80,0.5), seq(60,80,0.5), z, phi=35, 
        theta=35, col = color[facetcol],
        xlab="Father height", ylab="Mother height",zlab="Child height",
        main="Child Height Disregarding the Gender",
        ticktype="detailed")
persp(seq(60,80,0.5), seq(60,80,0.5), zMale, phi=35, 
        theta=35, col = color2[facetcol2],
        xlab="Father height", ylab="Mother height",zlab="Male height",
        main="Malec hild Height",
        ticktype="detailed" )
persp(seq(60,80,0.5), seq(60,80,0.5), zFemale, phi=35, 
        theta=35,  col = color3[facetcol3],
        xlab="Father height", ylab="Mother height",zlab="female height",
        main="Female child height" ,
        ticktype="detailed" )
```

By splitting the dataset by gender and plotting their graphs, it is possible to conclude that the child gender has some correlation with its height.

---
## Considering child gender

The application can show the difference between male and female height.
```{r}
fit <- expand.grid(list(x=68.46, y=63))
male <- predict(gfMale.loess, newdata=fit) 
female <- predict(gfFemale.loess, newdata=fit)
```
```{r}
male
female
```


---
## Conclusion
The child gender is a important estimator and can improve the prediction. In this application the LOESS method is used. However we don't mean this is the better algorithm for this problem.

### Thank you!



